Probabilistic Programming around Gaussian Processes

### 2019-10-21 abstract/paper.tex
lafi abstract in progress

\documentclass[sigplan,review]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
\usepackage{url}
\usepackage{hyperref}

\linespread{1.03}

@@

In this work, we advocate a third approach, which combines
advantages of the other two above. We present an architecture of
a Gaussian process library that provides an efficient
implementation of Gaussian process-related computations shared
by all algorithms, but lets program, virtually without
restrictions, both Gaussian process kernels and probabilistic
models of uncertainty, and combine Gaussian processes and
probabilistic models for inference using algorithms supported by
the framework. Our implementation of probabilistic
programming-enabled Gaussian processes, GoGP, is written in Go
and makes extensive use of modelling and inference capabilities
provided by Infergo~\cite{}, a Go library for differentiable
probabilistic programming. In what follows we first outline GoGP
design  and then illustrate on case studies, the use of GoGP in
a number of case studies inspired by applications.

@@

>>>>

\documentclass[sigplan,review]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
\usepackage{url}
\usepackage{hyperref}
\usepackage{listings}[language=Golang]
\usepackage{GoMono}

\lstdefinelanguage{Golang}%
  {morekeywords=[1]{package,import,func,type,struct,return,defer,panic,%
     recover,select,var,const,iota,},%
   morekeywords=[2]{string,uint,uint8,uint16,uint32,uint64,int,int8,int16,%
     int32,int64,bool,float32,float64,complex64,complex128,byte,rune,uintptr,%
     error,interface},%
   morekeywords=[3]{map,slice,make,new,nil,len,cap,copy,close,true,false,%
     delete,append,real,imag,complex,chan,},%
   morekeywords=[4]{for,break,continue,range,go,goto,switch,case,fallthrough,if,%
     else,default,},%
   morekeywords=[5]{Println,Printf,Error,Print,},%
   sensitive=true,%
   morecomment=[l]{//},%
   morecomment=[s]{/*}{*/},%
   morestring=[b]',%
   morestring=[b]",%
   morestring=[s]{`}{`},%
}

@@

In this work, we advocate a third approach, which combines
advantages of the other two above. We present an architecture of
a Gaussian process library that provides an efficient
implementation of Gaussian process-related computations shared
by all algorithms, but lets program, virtually without
restrictions, both Gaussian process kernels and probabilistic
models of uncertainty, and combine Gaussian processes and
probabilistic models for inference using algorithms supported by
the framework. Our implementation of probabilistic
programming-enabled Gaussian processes, GoGP, is written in Go
and makes extensive use of modelling and inference capabilities
provided by Infergo~\cite{}, a Go library for differentiable
probabilistic programming. In what follows we first outline GoGP
design  and then illustrate the use of GoGP on a number of case
studies inspired by applications.

@@

GoGP exports a type, \lstinline{GP}, implementing computations
involved in any use of a Gaussian process. The type has four
methods:

\begin{itemize}
	\item \lstinline{Absorb} --- adds (absorbs) observations;
	\item \lstinline{Produce} --- predicts out distributions;
	\item \lstlinline{Observe} --- computes log-likelihood of
		kernel hyperparameters given observations;
	\item \lstinline{Gradient} --- returns the gradient of
		log-likelihood by hyperparameters, and optionally
		by observations.
\end{itemize}

Methods \lstinline{Observe} and \lstinline{Gradient} also turn
\lstinline{GP} into an Infergo model with a custom gradient,
suitable for passing to Infergo inference algorithms.

package gp // import "bitbucket.org/dtolpin/gogp/gp"

TYPES

type GP struct {
	NDim                   int       // dimensions
	Simil, Noise           Kernel    // kernel
	ThetaSimil, ThetaNoise []float64 // kernel parameters

	// inputs
	X [][]float64 // inputs, for computing covariances
	Y []float64   // outputs

	// Has unexported fields.
}
    Type GP is the barebone implementation of GP.

func (gp *GP) Absorb(x [][]float64, y []float64) (err error)
    Absorb absorbs observations into the process.

func (gp *GP) Gradient() []float64
    Gradient computes the gradient of the log-likelihood with respect to the
    parameters and the inputs (GPML:5.9):

    ∇L = ½ tr((α α^⊤ - Σ^−1) ∂Σ/∂θ), where α = Σ^-1 y

func (gp *GP) LML() float64
    LML computes log marginal likelihood of the kernel given the absorbed
    observations (GPML:5.8):

    L = −½ log|Σ| − ½ y^⊤ α − n/2 log(2π), where α = Σ^-1 y

func (gp *GP) Observe(x []float64) float64
    Observe computes log marginal likelihood of the parameters given the
    observations. The argument is contatenation of log-transformed
    hyperparameters, inputs, and outputs.

    Optionally, the input can be only log-transformed hyperparameters, and then
    * only hyperparameters are inferred; * inputs must be assigned to fields X,
    Y of gp.

func (gp *GP) Produce(x [][]float64) (
	mu, sigma []float64,
	err error,
)
    Produce computes predictions.

type Kernel interface {
	model.Model
	NTheta() int
}
    Type Kernel is the kernel interface, implemented by both covariance and
    noise kernels.

GoGP
